Reverse Physics

### 2021-11-08 Articles/2021-ReversePhysics/ReversePhysics.tex
ReversePhysics: review before arxiv

\begin{abstract}
We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics. The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program. This allows us to identify physical starting points (as opposed to mathematical ones), compare their predictive power, and find those that can better serve as a foundation for our theories.
\end{abstract}

@@

Within the foundations of mathematics, the reverse mathematics program aims to start from known theorems and determine what axioms are required to prove them. This helps understand the relative logical strength of the theorems and get a better sense of their relationships. We propose a similar program for physics. We start from known physical theories or elements of physical theories, and we try to find premises that enable us to rederive them. The goal is to find premises that are physically more intuitive, so that ideally the arguments can be carried out with or without the math, and develop concepts that are of physical significance across theories.

While one does find efforts that analyze the premises of physical theories, for example in the context of foundations of quantum mechanics, statistical mechanics or the search for quantum gravity, their aim is typically different. Some are focused on a particular field (e.g. quantum theory) or try to demonstrate the centrality of one aspect (e.g. information) with regard to others. Some concentrate more on mathematical premises or structure, making the physical significance less than clear. Many aim to develop new theories, instead of first gaining further insight within current ones. We propose a more generalist approach that examines the relationships of physical ideas across subfields, which is best explained through examples.

To give a sense of the breadth of the reverse physics approach, we show how it can be used to pursue three different objectives. First, we illustrate multiple instances of reformulation. We study Hamilton's equation for a single degree of freedom to find six alternative characterizations that will connect different ideas from different fields under the same theme of determinism and reversibility. Next, we illustrate dependence analysis. We show that the uncertainty principle is not a consequence of the full quantum theory, but simply of its zero entropy bound of pure states. If a similar bound is imposed in classical mechanics, similar uncertainty relationships are recovered. Lastly, we illustrate reconceptualization. We take the third law of thermodynamics and rework it so that the role of crystalline substances to define zero entropy is instead played by the ``empty system''.

We want to stress that this paper is not about the originality of each single result, but about how these are used to solve a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.

@@

Let us group the state variables $\xi^a = \{q, p\}$ and consider the displacement vector field $S = \frac{d\xi^a}{dt} = \{ \frac{dq}{dt}, \frac{dp}{dt} \} $ which tells the direction in phase space in which each state moves over time. We find that $S$ is divergenceless:
\begin{align}\label{divergenceless}
\frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} =0
\end{align}
meaning that the flow through any closed region in phase space is zero. We also find the converse is true: a two dimensional divergenceless field $S$ always allows a stream function $H$ such that
\begin{align}
	S^q = \frac{\partial H}{\partial p} \; \; \; \; \; \; S^p = - \frac{\partial H}{\partial q}
\end{align}
In other words, Hamiltonian evolution for one degree of freedom is exactly the evolution for which the flow over a closed region is zero.

@@

It may seem odd that conservation of entropy for a deterministic system gives us energy conservation: we typically associate energy conservation with system isolation in thermodynamics. Where is the connection? For a deterministic and reversible system, future and past states depend only on the state of the present system. Therefore they do not depend on the state of other systems or of the environment. This means that deterministic and reversible systems are necessarily isolated. This time, running the argument in the opposite direction presents a problem. One needs to first decide whether a non-deterministic isolated system makes sense. If one assumes that the source of non-determinism, of stochastic uncertainty, is always interaction with other systems, other degrees of freedom, than every isolated system is deterministic. In this case, we can run the argument in reverse: every isolated system is deterministic and reversible. Without that extra assumption, this reverse direction is not guaranteed.

@@

We could go on to find other relationships and extend the ones we found to multiple degrees of freedom, but we believe this should be enough to highlight the power of the reverse physics approach. The first thing to note is how in a couple of pages we have found fundamental connections between classical Hamiltonian mechanics \eqref{HamiltonEquations1}, vector calculus \eqref{divergenceless}, differential geometry \eqref{UnitaryJacobian}, statistical mechanics \eqref{DeterminismReversibility}, thermodynamics \eqref{DeterminismThermodynamicReversibility}, information theory \eqref{EntropyConservation} and plain statistics \eqref{BijectionUncertainty}. This helps foster a sense of unity of these disparate disciplines and their perspectives, a sense that is sorely lacking both in research and education. We believe that a single unified view of this kind will bring more coherence and clarity to physics than, for example, a unified theory for the fundamental forces: nature is one, and does not care about how we have divided academic knowledge.

Another advantage is that we were able to carry out many of the arguments conceptually. Mathematics is then used to better articulate the physical arguments. This should please those who believe that physics, not mathematics, should be more at the center of the discussion.

@@

Let us now turn our attention to quantum mechanics, and see if our approach can shed new light on a theory that is not generally considered to be well understood. Since we talked about the role of uncertainty in classical mechanics, let us concentrate on the uncertainty principle which states that every state has to satisfy the relationship
\begin{align}\label{UncertaintyPrinciple}
	\sigma_q \sigma_p \geq \frac{\hbar}{2}.
\end{align}

We can start with an interesting observation. If we look back at \eqref{CovarianceMatrix}, we have:
\begin{align}
	\sigma_q^2 \sigma_p^2 = |\Sigma| + cov_{q,p}^2 \geq |\Sigma|.
\end{align}
Since by \eqref{BijectionUncertainty} $|\Sigma|$ is a constant of motion, we find that during Hamiltonian evolution the product of the uncertainty is bounded. The lowest uncertainty is reached when there are no correlations and the covariance is zero. Though it is a conceptually different relation, since $|\Sigma|$ is just a constant of motion, it gives us the following intuition: a coordinate invariant cap to the uncertainty will produce an inequality on the product of variances. Where can we find such a cap in quantum mechanics?

@@

We can test that hypothesis by studying the space of classical distributions with a fixed value of entropy $I_0$. We find that they have to satisfy the relationship\footnote{The careful reader will note the units do not quite work. The issue here is the $\log \rho$ in the Shannon/Gibbs entropy, as $\rho$ is not a pure number. Introducing a dimensionful constant $\log \rho \hat{h}$ would fix the expression, which would then fix the uncertainty relationship as well. } (TODO)
\begin{align}\label{ClassicalUncertaintyPrinciple}
	\sigma_q \sigma_p \geq \frac{e^{I_0}}{2 \pi e}.
\end{align}
The equality holds if the distribution is the product of two independent Gaussians.

The following condition:
\begin{align}\label{BoundedEntropy}
	\text{The entropy of the system is fixed}
\end{align}
is therefore enough to recover an uncertainty relationship independently of the theory. This can be understood as fixing the amount of information carried by the state. The reverse argument does not quite work, so condition \eqref{BoundedEntropy} is a stronger condition than the existence of an uncertainty relationship.

@@

In the previous section we saw the lower entropy bound of zero entropy is built into quantum mechanics and not in classical mechanics. However, the bound is also built into thermodynamics through the third law\footnote{This formulation is given by Gilbert Newton Lewis and Merle Randall, Thermodynamics and the Free Energy of
	Chemical Substances. McGraw-Hill, New York, 1923}
\begin{align}\label{ThirdLaw}
	\parbox{2.8in}{Every substance has a finite positive entropy, but
		at the absolute zero of temperature the entropy may become zero, and does so
		become in the case of perfect crystalline substances.}
\end{align}
Therefore we have two fundamental physical theories that share this trait. Can we find a more general characterization for this?

@@

Now that we have a better concept for zero entropy, we need to understand why we can't have states with lower entropy. Let us consider it from an information theoretical perspective. Suppose we had a state with negative entropy. This would mean that it is better specified than a state with zero entropy. The state is better specified than saying that the system is not there. But stating that the system is not there already completely specifies the system. Therefore the condition that imposes a lower bound of zero for entropy is the following:
\begin{align}\label{NewThirdLaw}
	\parbox{2.8in}{No state can describe a system more accurately than stating the system is not there in the first place.}
\end{align}
Again, note that this condition is not phenomenological: it is a logical necessity. This condition is realized in quantum mechanics: pure states represent the most precise descriptions of the system and they have zero entropy much like the vacuum, another pure state. This condition is realized in thermodynamics through the third law. In fact, the third law could be rederived as follows. Assuming \eqref{NewThirdLaw} tells us that entropy is bounded at zero. Since the entropy is a convex function of energy, entropy will reach its minimum at zero temperature, which is the only temperature that can reach zero entropy. We then use the previous result that the entropy of a crystalline substance at zero temperature is the same of an empty system.

@@

More than the particular results presented in this paper, it is the type of approach that we want to promote. We hope others will present their own reverse physics results, even better than those presented here, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.

\section{Proofs and calculations}

We omitted all proofs from the main body of the text because, as we stated, the physics should be given full attention. However, we do have to show that our arguments are on solid ground, so we include here basic proofs for all our major claims. To keep the discussion light, we will use the standards of rigor typically employed in physics. Note that most of the derivations are trivial, some already included in many textbooks. However, given the interdisciplinary nature of our subject, we find it useful for the reader to collect them all here.

\begin{prop}
	A dynamical system characterized by two variables $\{q, p\}$ follows Hamilton's equations if and only if the displacement field $S=\{\frac{dq}{dt}, \frac{dq}{dt} \}$ is divergenceless.
\end{prop}

\begin{proof}
	Suppose the evolution is Hamiltonian. Then $S=\{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$. The divergence is $div(S) = \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} = \frac{\partial \partial H}{\partial q\partial p} - \frac{\partial \partial H}{\partial p\partial q} = 0$.

	Conversely, suppose $S$ is divergenceless. Then, it admits a stream function $H$ such that $\{\frac{dq}{dt}, \frac{dq}{dt} = \{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$, which are Hamilton's equations
\end{proof}

\begin{prop}
	The displacement field $S=\{\frac{dq}{dt}, \frac{dq}{dt} \}$ of a dynamical system characterized by two variables $\{q, p\}$ is divergenceless if and only if the Jacobian determinant of the infinitesimal time evolution is equal to one.
\end{prop}

@@

\begin{remark}
	An unstated assumption here is that $H$ is twice differentiable. The Jacobian and the divergence, in fact, would not be well defined if $H$ were differentiable only. In that case, the acceleration $a = \frac{d}{dt}\frac{dq}{dt} = \frac{d}{dt} \frac{\partial H}{\partial p}$ would also be ill defined. With this in mind, if a Hamiltonian is not twice differentiable at a point it is better to consider the system not Hamiltonian at that point.
\end{remark}

@@

\begin{proof}
	Let $I[\rho(x^i)] = - \int \rho(x^i) \log \rho(x^i) \, dx^n$ be the Shannon entropy of the distribution $\rho(x^i)$. Let $y^j=y^j(x^i)$ be a differentiable transformation. Since $\rho(x^i)$ is a density, we have $\rho(y^j) = \rho(x^i) |J|$. The information entropy after the transformation will be
	\begin{align*}
		I[\rho(y^j)] &= - \int \rho(y^j) \log \rho(y^j) \, dy^n \\
		&= - \int \rho(x^i) |J| \log \left( \rho(x^i) |J| \right) \, dy^n \\
		&= - \int \rho(x^i) \log \left( \rho(x^i) |J| \right) \, dx^n \\
		&= - \int \rho(x^i) \log \rho(x^i) \, dx^n - \int \rho(x^i) \log |J| \, dx^n \\
		&= I[\rho(x^i)] - \int \rho(x^i) \log |J| \, dx^n.
	\end{align*}
	The information entropy to be conserved for every $\rho$ exactly when $|J|$ is $\pm 1$ everywhere. Given that the transformation is continuous, the Jacobian determinant must be positive (i.e. a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
\end{proof}

@@

\begin{proof}
	As we assume the distribution to be small enough, we can linearize the transformation to $y^j = J^j_i x^i + B^j$ where $J$ is the Jacobian. Noting that the covariance is a linear operator that does not depend on the expectation of the variables, we have
	\begin{align*}
		|\Sigma^{ij}| &= |cov(y^i, y^j)| \\
		&= |cov(J^i_k x^k + B^i, J^j_l x^l + B^j)| \\
		&= |cov(J^i_k x^k , J^j_l x^l )| \\
		&= |J^i_k cov(x^k , x^l ) J^j_l| \\
		&= |J^i_k | | \Sigma^{kl} | | J^j_l| \\
	\end{align*}
	The uncertainty is conserved for all small distributions exactly when $|J|$ is $\pm 1$ everywhere. Given that the transformation is continuous, the Jacobian determinant must be positive (i.e. a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
\end{proof}

@@

\begin{proof}
	Let $\rho$ be the density matrix of a pure state. We have $\rho = |\psi \rangle \langle psi | = |\psi \rangle \langle psi |\psi \rangle \langle psi | = \rho^2$. The von Neumann entropy is $I[\rho] = - tr (\rho \log \rho) = - tr (\rho \log \rho^2) = - 2 tr (\rho \log \rho) = 2 I[\rho]$. Therefore $I[\rho] = 0$.
\end{proof}

@@

\begin{proof}

\end{proof}

>>>>

\begin{abstract}
We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics. The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program. This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
\end{abstract}

@@

Within the foundations of mathematics, the reverse mathematics program\cite{friedman1976systems,simpson2017reverse,stillwellreverse} aims to start from known theorems and determine what axioms are required to prove them. This helps understand the relative logical strength of the theorems and get a better sense of their relationships. We propose a similar program for physics. We start from known physical theories or elements of physical theories, and we try to find premises that enable us to rederive them. The goal is to find premises that are physically more intuitive, so that ideally the arguments can be carried out with or without the math, and develop concepts that are of physical significance across theories.

While one does find efforts that analyze the premises of physical theories (see e.g. Refs.~\cite{chiribella2011informational, selby2021reconstructing, giles2016mathematical, boyling1972axiomatic, PhysRev.101.860, Haag1964848, axiomaticQFT1975, masanes2019measurement, carcassi2021four}), for example in the context of foundations of quantum mechanics, statistical mechanics or the search for quantum gravity, their aim is typically different. Some are focused on a particular field (e.g. quantum theory) or try to demonstrate the centrality of one aspect (e.g. information) with regard to others. Some concentrate more on mathematical premises or structure, making the physical significance less than clear. Many aim to develop new theories, instead of first gaining further insight within current ones. We propose a more generalist approach that examines the relationships of physical ideas across subfields, which is best explained through examples.

To give a sense of the breadth of the reverse physics approach, we show how it can be used to pursue three different objectives. First, we illustrate multiple instances of \emph{reformulation}. We study Hamilton's equations for a single degree of freedom to find six alternative characterizations that will connect different ideas from different fields under the same theme of determinism and reversibility. Next, we illustrate \emph{dependence analysis}. We show that the uncertainty principle is not a consequence of the full quantum theory, but simply of its zero entropy bound on pure states. If a similar bound is imposed in classical mechanics, similar uncertainty relationships are recovered. Lastly, we illustrate \emph{reconceptualization}. We take the third law of thermodynamics and rework it so that the role of crystalline substances to define zero entropy is instead played by the ``empty system''.

As we want the discussion to be focused on the physics, we will use the most widely known math among physicists and relegate mathematical derivations to the appendix. We also want to stress that this paper is not about the originality of each single result, but about how these are used to address a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.

@@

Let us group the state variables $\xi^a = \{q, p\}$ and consider the displacement vector field $S = \frac{d\xi^a}{dt} = \{ \frac{dq}{dt}, \frac{dp}{dt} \} $ which tells the direction in phase space in which each state moves over time. We find that $S$ is divergenceless:
\begin{align}\label{divergenceless}
\frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} =0
\end{align}
meaning that the flow through any closed region in phase space is zero. We also find the converse is true: a two dimensional divergenceless field $S$ always allows a stream function $H$ such that
\begin{align}
	S^q = \frac{\partial H}{\partial p} \; \; \; \; \; \; S^p = - \frac{\partial H}{\partial q}
\end{align}
In other words, Hamiltonian evolution for one degree of freedom is exactly the evolution for which the flow over a closed region in phase space is zero.

@@

It may seem odd that conservation of entropy for a deterministic system gives us energy conservation: we typically associate energy conservation with system isolation in thermodynamics. Where is the connection? For a deterministic and reversible system, future and past states depend only on the state of the present system. Therefore they do not depend on the state of other systems or of the environment. This means that deterministic and reversible systems are necessarily isolated. This time, running the argument in the opposite direction presents a problem. One needs to first decide whether a non-deterministic isolated system makes sense. If one assumes that the source of non-determinism, of stochastic uncertainty, is always interaction with other systems, other degrees of freedom, then every isolated system is deterministic. In this case, we can run the argument in reverse: every isolated system is deterministic and reversible. Without that extra assumption, this reverse direction is not guaranteed.

@@

We could go on to find other relationships and extend the ones we found to multiple degrees of freedom, but we believe this should be enough to highlight the power of the reverse physics approach. The first thing to note is how in a couple of pages we have found fundamental connections between classical Hamiltonian mechanics \eqref{HamiltonEquations1}, vector calculus \eqref{divergenceless}, differential geometry \eqref{UnitaryJacobian}, statistical mechanics \eqref{DeterminismReversibility}, thermodynamics \eqref{DeterminismThermodynamicReversibility}, information theory \eqref{EntropyConservation} and plain statistics \eqref{BijectionUncertainty}. This helps foster a sense of unity of these disparate disciplines and their perspectives, a sense that is sorely lacking both in research and education: nature is one, and does not care about how we have divided academic knowledge. We believe that a single unified view of this kind will bring more coherence and clarity to physics than, for example, a unified theory for the fundamental forces.

Another advantage is that we were able to carry out many of the arguments conceptually. Mathematics is then used to better articulate the physical arguments. This should please those who believe that physics, not mathematics, should be more at the center of the discussion.\cite{hossenfelder2018lost, woit2006not}

@@

Let us now turn our attention to quantum mechanics, and see if our approach can shed new light on a theory that is not generally considered to be well understood. Since we talked about the role of uncertainty in classical mechanics, let us concentrate on the uncertainty principle, which states that every state has to satisfy the relationship
\begin{align}\label{UncertaintyPrinciple}
	\sigma_q \sigma_p \geq \frac{\hbar}{2}.
\end{align}

We can start with an interesting observation. If we look back at \eqref{CovarianceMatrix}, we have:
\begin{align}
	\sigma_q^2 \sigma_p^2 = |\Sigma| + cov_{q,p}^2 \geq |\Sigma|.
\end{align}
Since by \eqref{BijectionUncertainty} $|\Sigma|$ is a constant of motion, we find that during Hamiltonian evolution the uncertainty is bounded. The lowest uncertainty is reached when there are no correlations and the covariance is zero. Though it is a conceptually different relation, since $|\Sigma|$ is just a constant of motion, it gives us the following intuition: a coordinate invariant cap to the uncertainty will produce an inequality on the product of variances. Where can we find such a cap in quantum mechanics?

@@

We can test that hypothesis by studying the space of classical distributions with a fixed value of entropy $I_0$. We find that they have to satisfy the relationship\footnote{The careful reader will note the units do not quite work. The issue here is the $\log \rho$ in the Shannon/Gibbs entropy, as $\rho$ is not a pure number. Introducing a dimensionful constant $\log \rho \hat{h}$ would fix the expression, which would then fix the uncertainty relationship as well. }
\begin{align}\label{ClassicalUncertaintyPrinciple}
	\sigma_q \sigma_p \geq \frac{e^{I_0}}{2 \pi e}.
\end{align}
The equality holds if the distribution is the product of two independent Gaussians.

The following condition:
\begin{align}\label{BoundedEntropy}
	\text{The entropy of the system is fixed}
\end{align}
is therefore enough to recover an uncertainty relationship independently of the theory. This can be understood as fixing the amount of information carried by the state. The reverse argument does not work, so condition \eqref{BoundedEntropy} is a stronger condition than the existence of an uncertainty relationship.

@@

In the previous section we saw the lower entropy bound of zero entropy is built into quantum mechanics and not in classical mechanics. However, the bound is also built into thermodynamics through the third law\footnote{This formulation is given by Ref. \cite{lewis1923thermodynamics}}
\begin{align}\label{ThirdLaw}
	\parbox{2.8in}{Every substance has a finite positive entropy, but
		at the absolute zero of temperature the entropy may become zero, and does so
		become in the case of perfect crystalline substances.}
\end{align}
Therefore we have two fundamental physical theories that share this trait. Can we find a more general characterization for this?

@@

Now that we have a better concept for zero entropy, we need to understand why we can't have states with lower entropy. Let us consider it from an information theoretical perspective. Suppose we had a state with negative entropy. This would mean that it is better specified than a state with zero entropy. The state is better specified than saying that the system is not there. But stating that the system is not there already completely specifies the system. Therefore the condition that imposes a lower bound of zero for entropy is the following:
\begin{align}\label{NewThirdLaw}
	\parbox{2.8in}{No state can describe a system more accurately than stating the system is not there in the first place.}
\end{align}
Again, note that this condition is not phenomenological: it is a logical necessity. This condition is realized in quantum mechanics: pure states represent the most precise descriptions of the system and they have zero entropy much like the vacuum, another pure state. This condition is realized in thermodynamics through the third law. In fact, the third law could be rederived as follows. Assuming \eqref{NewThirdLaw} tells us that entropy is bounded at zero. Since the entropy in thermodynamics is a convex function of energy, entropy will reach its minimum at zero temperature, which is the only temperature that can reach zero entropy. We then use the previous result that the entropy of a crystalline substance at zero temperature is the same as for an empty system.

@@

More than the particular results presented in this paper, it is the type of approach that we want to promote. We hope others will present their own reverse physics results, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.

\section*{Appendix: Proofs and calculations}

We include here all proofs which we omitted from the main body since, as we stated, the physics should be given full attention. To keep the discussion light and accessible to the widest audience, we use the simplest mathematical techniques to get the result. Some readers may feel that some results are well known, or too simple to even be included. Due to the interdisciplinary nature of the subject, we believe it is useful to collect them all here.

\begin{prop}
	A dynamical system characterized by two variables $\{q, p\}$ follows Hamilton's equations if and only if the displacement field $S=\{\frac{dq}{dt}, \frac{dp}{dt} \}$ is divergenceless.
\end{prop}

\begin{proof}
	Suppose the evolution is Hamiltonian. Then $S=\{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$. The divergence is $div(S) = \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} = \frac{\partial^2 H}{\partial q\partial p} - \frac{\partial^2H}{\partial p\partial q} = 0$.

	Conversely, suppose $S$ is divergenceless. Then, it admits a stream function $H$ such that $\{\frac{dq}{dt}, \frac{dp}{dt}\} = \{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$, which are Hamilton's equations.
\end{proof}

\begin{prop}
	The displacement field $S=\{\frac{dq}{dt}, \frac{dp}{dt} \}$ of a dynamical system characterized by two variables $\{q, p\}$ is divergenceless if and only if the Jacobian determinant of the infinitesimal time evolution is equal to one.
\end{prop}

@@

\begin{remark}
	An unstated assumption here is that $H$ is twice differentiable. The Jacobian and the divergence, in fact, would not be well defined if $H$ were differentiable only once. In that case, the acceleration $a = \frac{d}{dt}\frac{dq}{dt} = \frac{d}{dt} \frac{\partial H}{\partial p}$ would also be ill defined. With this in mind, if a Hamiltonian is not twice differentiable at a point it is better to consider the system not Hamiltonian at that point.
\end{remark}

@@

\begin{proof}
	Let $I[\rho(x^i)] = - \int \rho(x^i) \log \rho(x^i) \, dx^n$ be the Shannon entropy of the distribution $\rho(x^i)$. Let $y^j=y^j(x^i)$ be a differentiable transformation. Since $\rho(x^i)$ is a density, we have $\rho(y^j) = \rho(x^i) |J|$. The information entropy after the transformation will be
	\begin{align*}
		I[\rho(y^j)] &= - \int \rho(y^j) \log \rho(y^j) \, dy^n \\
		&= - \int \rho(x^i) |J| \log \left( \rho(x^i) |J| \right) \, dy^n \\
		&= - \int \rho(x^i) \log \left( \rho(x^i) |J| \right) \, dx^n \\
		&= - \int \rho(x^i) \log \rho(x^i) \, dx^n - \int \rho(x^i) \log |J| \, dx^n \\
		&= I[\rho(x^i)] - \int \rho(x^i) \log |J| \, dx^n.
	\end{align*}
	The information entropy is conserved for every $\rho$ if and only if $|J|$ is 1 everywhere. Therefore the entropy is conserved if and only if the Jacobian is unitary.
\end{proof}

@@

\begin{proof}
	As we assume the distribution to be small enough, we can linearize the transformation to $y^j = J^j_i x^i + B^j$ where $J$ is the Jacobian. Noting that the covariance is a linear operator that does not depend on the expectation of the variables, we have
	\begin{align*}
		|\Sigma^{ij}| &= |cov(y^i, y^j)| \\
		&= |cov(J^i_k x^k + B^i, J^j_l x^l + B^j)| \\
		&= |cov(J^i_k x^k , J^j_l x^l )| \\
		&= |J^i_k cov(x^k , x^l ) J^j_l| \\
		&= |J^i_k | | \Sigma^{kl} | | J^j_l| \\
	\end{align*}
	The uncertainty is conserved for all small distributions if and only if $|J|$ is $\pm 1$ everywhere. Given that the transformation is continuous, the Jacobian determinant must be positive (i.e. a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
\end{proof}

@@

\begin{proof}
	Let $\rho$ be the density matrix of a pure state. We have $\rho = |\psi \rangle \langle \psi | = |\psi \rangle \langle \psi |\psi \rangle \langle \psi | = \rho^2$. The von Neumann entropy is $I[\rho] = - tr (\rho \log \rho) = - tr (\rho \log \rho^2) = - 2 tr (\rho \log \rho) = 2 I[\rho]$. Therefore $I[\rho] = 0$.
\end{proof}

@@

\begin{proof}
	We set up a minimization problem using Lagrange multipliers. We want to minimize the product of the variance $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$ while keeping $\rho$ normalized and fixing the entropy to $I_0$. We have: 
	\begin{align*}
		L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
		&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) \\ &+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
		\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
		\lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
		\rho_{\mathcal{c}} = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}
	\end{align*}
	We solve the multipliers and have:
	\begin{align*}
		\rho_{\mathcal{c}} = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
		I_0 = &\ln (2\pi e\sigma_q\sigma_p) \\
		\sigma_q \sigma_p &= \frac{e^{I_0}}{2 \pi e}
	\end{align*}
	This shows that the gaussian minimizes the spread at fixed entropy, therefore all other distributions must have a larger or equal spread. 
\end{proof}

\begin{remark}
	Note that the inverse argument does not work: if we have a bound on the uncertainty, we cannot say anything about the entropy. All distributions with higher entropy will satisfy a higher bound, therefore the entropy is arbitrarily high. Moreover, we can find distributions with low entropy but with arbitrarily high spread. Therefore a bound on the uncertainty still allows any value for entropy.
\end{remark}
